Model_CFU

### 2017-01-05 code/HumanCdGF_cdiff_colonization.Rmd
add baseline/randomized models

set.seed(seed)
cfu_all_rf <- randomForest(cfu_all_shared_df[ , -1], cfu_all_shared_df$log_cfu_med,
                           ntree = 1001,
                           mtry = 15,
                           importance = T)
top_features <- head(sort(importance(cfu_all_rf)[,1],decreasing = T), n_features)
feat_cor_cfu <- cor(cfu_all_shared_df[ , c('log_cfu_med', 
                                           names(top_features))],
                    method = 'spearman')[-1,'log_cfu_med']
```

@@

data.frame(Observed = cfu_all_shared_df$log_cfu_med, 
           Predicted = cfu_all_rf$predicted) %>% 
  ggplot(aes(x = Predicted, y = Observed)) + geom_point() +
    xlim(5,9) + ylim(5,9) + 
    annotate('text', label = paste0(round(tail(cfu_all_rf$rsq, 1), 4)*100, '% Var Explained'),
             x = min(cfu_all_rf$predicted), y = max(cfu_all_shared_df$log_cfu_med)) + 
    theme(panel.background = element_rect(linetype = 1, size = 1, fill = 'white', color = 'black'),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank()) +
  labs(x = 'Predicted Colonization (log CFU)', y = 'Observed Colonization (log CFU)')

@@

@@

set.seed(seed)
source_all_rf <- randomForest(source_all_shared_df[ , -1], source_all_shared_df$human_source,
                           ntree = 501,
                           mtry = 15,
                           importance = T)
top_features_src <- head(sort(importance(source_all_rf)[,1],decreasing = T), n_features)
#any(names(top_features) %in% names(top_features_src))
src_feat_cor_cfu <- cor(cfu_all_shared_df[ , c('log_cfu_med', 
                                               names(top_features_src))], 
                        method = 'spearman')[-1,'log_cfu_med']
```

@@

@@

set.seed(seed)
cfu_all_src_rf <- randomForest(cfu_all_src_shared_df[ , -1], cfu_all_src_shared_df$log_cfu_med,
                           ntree = 1001,
                           mtry = 15,
                           importance = T)
top_features_src_cfu <- head(sort(importance(cfu_all_src_rf)[,1],decreasing = T), n_features)
feature_df_src_numeric <- cfu_all_src_shared_df[ , c('log_cfu_med', names(top_features_src_cfu))]
feature_df_src_numeric$human_source <- as.numeric(feature_df_src_numeric$human_source)
feat_cor_cfu_src_all <- cor(feature_df_src_numeric, method = 'spearman')[-1,'log_cfu_med']
```

@@

data.frame(Observed = cfu_all_src_shared_df$log_cfu_med, 
           Predicted = cfu_all_src_rf$predicted) %>% 
  ggplot(aes(x = Predicted, y = Observed)) + geom_point() +
    xlim(5,9) + ylim(5,9) + 
    annotate('text', label = paste0(round(tail(cfu_all_src_rf$rsq, 1), 4)*100, '% Var Explained'),
             x = min(cfu_all_src_rf$predicted), y = max(cfu_all_src_shared_df$log_cfu_med)) + 
    theme(panel.background = element_rect(linetype = 1, size = 1, fill = 'white', color = 'black'),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank()) +
  labs(x = 'Predicted Colonization (log CFU)', y = 'Observed Colonization (log CFU)')

@@

@@

set.seed(seed)
cfu_wo581_rf <- randomForest(cfu_wo581_df[ , -2], cfu_wo581_df$log_cfu_med,
                           ntree = 1001,
                           mtry = 15,
                           importance = T)
top_features_wo581 <- head(sort(importance(cfu_wo581_rf)[,1],decreasing = T), n_features)

@@

```{r colonization performance without 581, fig.width=5, fig.height=4}
data.frame(Observed = cfu_wo581_df$log_cfu_med, 
           Predicted = cfu_wo581_rf$predicted) %>% 
  ggplot(aes(x = Predicted, y = Observed)) + geom_point() +
    xlim(5,9) + ylim(5,9) + 
    annotate('text', label = paste0(round(tail(cfu_wo581_rf$rsq, 1), 4)*100, '% Var Explained'),
             x = 6, y = 8.9) + 
    theme(panel.background = element_rect(linetype = 1, size = 1, fill = 'white', color = 'black'),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank()) +
  labs(x = 'Predicted Colonization (log CFU)', y = 'Observed Colonization (log CFU)')

@@

@@

```{r all without 581 or 1134/1146}
cfu_wo_hilo_df <- colonized_cfu %>% 
    select(mouse_id, human_source, log_cfu_med) %>% 
    right_join(community_all_d0) %>% 
    filter(!human_source %in% c('DA00581', 'DA01134','DA01146')) %>% 
    select(-mouse_id) %>% 
    filter(!is.na(log_cfu_med))
cfu_wo_hilo_df$human_source <- factor(cfu_wo_hilo_df$human_source)

set.seed(seed)
cfu_wo_hilo_rf <- randomForest(cfu_wo_hilo_df[ , -2], cfu_wo_hilo_df$log_cfu_med,
                           ntree = 1001,
                           mtry = 15,
                           importance = T)
top_features_wo_hilo <- head(sort(importance(cfu_wo_hilo_rf)[,1],decreasing = T), n_features)

data.frame(Observed = cfu_wo_hilo_df$log_cfu_med, 
           Predicted = cfu_wo_hilo_rf$predicted) %>% 
  ggplot(aes(x = Predicted, y = Observed)) + geom_point() +
    xlim(5,9) + ylim(5,9) + 
    annotate('text', label = paste0(round(tail(cfu_wo_hilo_rf$rsq, 1), 4)*100, '% Var Explained'),
             x = 5.5, y = 8.9) + 
    theme(panel.background = element_rect(linetype = 1, size = 1, fill = 'white', color = 'black'),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank()) +
  labs(x = 'Predicted Colonization (log CFU)', y = 'Observed Colonization (log CFU)')
```

>>>>

baseline_df <- c()
baseline_rsq <- c()
for(i in 1:100){
  set.seed(i)
  random_cfu <- sample(cfu_all_shared_df$log_cfu_med, 
                       length(cfu_all_shared_df$log_cfu_med), replace = F)
  set.seed(seed)
  baseline_cfu_all_rf <- randomForest(cfu_all_shared_df[ , -1], random_cfu,
                           ntree = 501,
                           mtry = 15)
  baseline_df <- rbind(baseline_df, data.frame(Model = 'Baseline', Observed = random_cfu,
                                               Predicted = baseline_cfu_all_rf$predicted))
  baseline_rsq <- c(baseline_rsq, tail(baseline_cfu_all_rf$rsq,1))
}
set.seed(seed)
cfu_all_rf <- randomForest(cfu_all_shared_df[ , -1], cfu_all_shared_df$log_cfu_med,
                           ntree = 501,
                           mtry = 15,
                           importance = T)
top_features <- head(sort(importance(cfu_all_rf)[,1],decreasing = T), n_features)
feat_cor_cfu <- cor(cfu_all_shared_df[ , c('log_cfu_med', 
                                           names(top_features))],
                    method = 'spearman')[-1,'log_cfu_med']
```

@@

rbind(baseline_df,
      data.frame(Model = 'Prediction', Observed = cfu_all_shared_df$log_cfu_med, 
           Predicted = cfu_all_rf$predicted)) %>% 
  ggplot(aes(x = Predicted, y = Observed, color=Model)) + geom_point() +
    xlim(5,9) + ylim(5,9) + scale_colour_manual(values = c('#CCCCCC','#000000')) +
    annotate('text', label = paste0(round(tail(cfu_all_rf$rsq, 1), 4)*100, '% Var Explained'),
             x = min(cfu_all_rf$predicted), y = max(cfu_all_shared_df$log_cfu_med)) + 
    theme(panel.background = element_rect(linetype = 1, size = 1, fill = 'white', color = 'black'),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank(),
          legend.position = c(0.9,0.1)) +
  labs(x = 'Predicted Colonization (log CFU)', y = 'Observed Colonization (log CFU)')

@@

The % Variation explained in the baseline model is:  
Min.  1st Qu.   Median     Mean  3rd Qu.     Max.   
`r summary(baseline_rsq)`

@@

baseline_source_df <- c()
for(i in 1:100){
  set.seed(i)
  random_src <- sample(source_all_shared_df$human_source, 
                       length(source_all_shared_df$human_source), replace = F)
  set.seed(seed)
  baseline_src_all_rf <- randomForest(source_all_shared_df[ , -1], random_src,
                           ntree = 501,
                           mtry = 15)
  baseline_source_df <- c(baseline_source_df, tail(baseline_src_all_rf$err.rate[,'OOB'],1))
}
set.seed(seed)
source_all_rf <- randomForest(source_all_shared_df[ , -1], source_all_shared_df$human_source,
                           ntree = 501,
                           mtry = 15,
                           importance = T)
top_features_src <- head(sort(importance(source_all_rf)[,1],decreasing = T), n_features)
#any(names(top_features) %in% names(top_features_src))
src_feat_cor_cfu <- cor(cfu_all_shared_df[ , c('log_cfu_med', 
                                               names(top_features_src))], 
                        method = 'spearman')[-1,'log_cfu_med']
```

@@

Models predicting randomized donor perform with a high error rate:   
Mean Error Rate = `r summary(baseline_source_df)[4]` with an IQR of `r summary(baseline_source_df)[2]` - `r summary(baseline_source_df)[5]`.

@@

baseline_cfu_src_df <- c()
baseline_src_rsq <- c()
for(i in 1:100){
  set.seed(i)
  random_cfu <- sample(cfu_all_src_shared_df$log_cfu_med, 
                       length(cfu_all_src_shared_df$log_cfu_med), replace = F)
  set.seed(seed)
  baseline_cfu_src_rf <- randomForest(cfu_all_src_shared_df[ , -1], random_cfu,
                           ntree = 501,
                           mtry = 15)
  baseline_cfu_src_df <- rbind(baseline_cfu_src_df, 
                               data.frame(Model = 'Baseline', Observed = random_cfu,
                                          Predicted = baseline_cfu_src_rf$predicted))
  baseline_src_rsq <- c(baseline_src_rsq, tail(baseline_cfu_src_rf$rsq,1))

}
set.seed(seed)
cfu_all_src_rf <- randomForest(cfu_all_src_shared_df[ , -1], cfu_all_src_shared_df$log_cfu_med,
                           ntree = 1001,
                           mtry = 15,
                           importance = T)
top_features_src_cfu <- head(sort(importance(cfu_all_src_rf)[,1],decreasing = T), n_features)
feature_df_src_numeric <- cfu_all_src_shared_df[ , c('log_cfu_med', names(top_features_src_cfu))]
feature_df_src_numeric$human_source <- as.numeric(feature_df_src_numeric$human_source)
feat_cor_cfu_src_all <- cor(feature_df_src_numeric, method = 'spearman')[-1,'log_cfu_med']
```

@@

rbind(baseline_cfu_src_df,
      data.frame(Model = 'Prediction',
                 Observed = cfu_all_src_shared_df$log_cfu_med, 
                 Predicted = cfu_all_src_rf$predicted)) %>% 
  ggplot(aes(x = Predicted, y = Observed, color = Model)) + geom_point() +
    xlim(5,9) + ylim(5,9) + scale_colour_manual(values = c('#CCCCCC','#000000')) +
    annotate('text', label = paste0(round(tail(cfu_all_src_rf$rsq, 1), 4)*100, '% Var Explained'),
             x = min(cfu_all_src_rf$predicted), y = max(cfu_all_src_shared_df$log_cfu_med)) + 
    theme(panel.background = element_rect(linetype = 1, size = 1, fill = 'white', color = 'black'),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank(),
          legend.position = c(0.9,0.1)) +
  labs(x = 'Predicted Colonization (log CFU)', y = 'Observed Colonization (log CFU)')

@@

The % Variation explained in the baseline model is:  
Min.  1st Qu.   Median     Mean  3rd Qu.     Max.   
`r summary(baseline_src_rsq)`

@@

baseline_wo581 <- c()
base_wo581_rsq <- c()
for(i in 1:100){
  set.seed(i)
  random_cfu <- sample(cfu_wo581_df$log_cfu_med, 
                       length(cfu_wo581_df$log_cfu_med), replace = F)
  set.seed(seed)
  baseline_wo581_rf <- randomForest(cfu_wo581_df[ , -1], random_cfu,
                           ntree = 501,
                           mtry = 15)
  baseline_wo581 <- rbind(baseline_wo581, 
                               data.frame(Model = 'Baseline', Observed = random_cfu,
                                          Predicted = baseline_wo581_rf$predicted))
  base_wo581_rsq <- c(base_wo581_rsq, tail(baseline_wo581_rf$rsq,1))
}
set.seed(seed)
cfu_wo581_rf <- randomForest(cfu_wo581_df[ , -2], cfu_wo581_df$log_cfu_med,
                           ntree = 1001,
                           mtry = 15,
                           importance = T)
top_features_wo581 <- head(sort(importance(cfu_wo581_rf)[,1],decreasing = T), n_features)

@@

```{r colonization performance without 581, fig.width=5, fig.height=4}
rbind(baseline_wo581,
      data.frame(Model = 'Prediction', Observed = cfu_wo581_df$log_cfu_med, 
           Predicted = cfu_wo581_rf$predicted)) %>% 
  ggplot(aes(x = Predicted, y = Observed, color = Model)) + geom_point() +
    xlim(5,9) + ylim(5,9) + scale_colour_manual(values = c('#CCCCCC','#000000')) +
    annotate('text', label = paste0(round(tail(cfu_wo581_rf$rsq, 1), 4)*100, '% Var Explained'),
             x = 6, y = 8.9) + 
    theme(panel.background = element_rect(linetype = 1, size = 1, fill = 'white', color = 'black'),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank(),
          legend.position = c(0.8,0.2)) +
  labs(x = 'Predicted Colonization (log CFU)', y = 'Observed Colonization (log CFU)')

@@

The % Variation explained in the baseline model is:  
Min.  1st Qu.   Median     Mean  3rd Qu.     Max.   
`r summary(base_wo581_rsq)`

@@

```{r all without 581 or 1134/1146, fig.width=5, fig.height=4}
cfu_wo_hilo_df <- colonized_cfu %>% 
    select(mouse_id, human_source, log_cfu_med) %>% 
    right_join(community_all_d0) %>% 
    filter(!human_source %in% c('DA00581', 'DA01134','DA01146')) %>% 
    select(-mouse_id) %>% 
    filter(!is.na(log_cfu_med))
cfu_wo_hilo_df$human_source <- factor(cfu_wo_hilo_df$human_source)

baseline_wo_hilo <- c()
base_wo_hilo_rsq <- c()
for(i in 1:100){
  set.seed(i)
  random_cfu <- sample(cfu_wo_hilo_df$log_cfu_med, 
                       length(cfu_wo_hilo_df$log_cfu_med), replace = F)
  set.seed(seed)
  baseline_wo_hilo_rf <- randomForest(cfu_wo_hilo_df[ , -1], random_cfu,
                           ntree = 501,
                           mtry = 15)
  baseline_wo_hilo <- rbind(baseline_wo_hilo, 
                               data.frame(Model = 'Baseline', Observed = random_cfu,
                                          Predicted = baseline_wo_hilo_rf$predicted))
  base_wo_hilo_rsq <- c(base_wo_hilo_rsq, tail(baseline_wo_hilo_rf$rsq,1))
}
set.seed(seed)
cfu_wo_hilo_rf <- randomForest(cfu_wo_hilo_df[ , -2], cfu_wo_hilo_df$log_cfu_med,
                           ntree = 1001,
                           mtry = 15,
                           importance = T)
top_features_wo_hilo <- head(sort(importance(cfu_wo_hilo_rf)[,1],decreasing = T), n_features)

rbind(baseline_wo_hilo,
      data.frame(Model = 'Prediction', Observed = cfu_wo_hilo_df$log_cfu_med, 
           Predicted = cfu_wo_hilo_rf$predicted)) %>% 
  ggplot(aes(x = Predicted, y = Observed, color = Model)) + geom_point() +
    xlim(5,9) + ylim(5,9) + scale_colour_manual(values = c('#CCCCCC','#000000')) +
    annotate('text', label = paste0(round(tail(cfu_wo_hilo_rf$rsq, 1), 4)*100, '% Var Explained'),
             x = 6, y = 8.9) + 
    theme(panel.background = element_rect(linetype = 1, size = 1, fill = 'white', color = 'black'),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank(),
          legend.position = c(0.8,0.2)) +
  labs(x = 'Predicted Colonization (log CFU)', y = 'Observed Colonization (log CFU)')
```

The % Variation explained in the baseline model is:  
Min.  1st Qu.   Median     Mean  3rd Qu.     Max.   
`r summary(base_wo_hilo_rsq)`
